---
title: "Assignment 5"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---

```{r setup, message=FALSE, echo=FALSE, warning=FALSE}
library(data.table)
library(tidyverse)
library(dplyr)
library(plotly)
library(DT)
library(knitr)

# INITALIZE CODE CHUNK OPTIONS
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  eval=TRUE,
  echo = FALSE,
  cache = TRUE,
  fig.width = 7, 
  fig.align = 'center',
  fig.asp = 0.618,
  out.width = "700px",
  class.source = "code-r")
```

```{css, echo = FALSE}
.code-r { /* Code block */
  font-size: 15px;
}

.code-r-small { /* Code block */
  font-size: 10px;
}
```

```{r}
library(skimr)
library(olsrr)
library(mfp)
```

# Assignment 5 Final Project Interactive Visualization

<br>

## Data description 

- Lipotoxicity: lipotoxicity is a metabolic syndrome that results from the accumulation of lipid intermediates in non-adipose tissue, leading to cellular dysfunction and death.

- Glucose: blood glucose level is the concentration of glucose present in the blood.

- BloodPressure: the pressure of the blood in the circulatory system.

- SkinThickness: skin thickness is primarily determined by collagen content and is increased in insulin-dependent diabetes mellitus (IDDM).

- Insulin: insulin is an anabolic hormone that promotes glucose uptake.

- BMI: body mass index (BMI) is a person's weight in kilograms divided by the square of height in meters.

- DiabetesPedigreeFunction: diabetes pedigree function provides “a synthesis of the diabetes mellitus history in relatives and the genetic relationship of those relatives to the subject.” It utilizes information from a person’s family history to predict how diabetes will affect that individual.

- Age: age of the individual.

- Outcome: diabetes diagnosis.

<br>

## Key question
- Whether diabetes risk factors are associated with the outcome. Whether Diabetes Pedigree Function has significant association with outcome and other risk factors. 

```{r}
diabetes <- read_csv("data/diabetes.csv")
```

```{r}
db <-
  diabetes %>%
  mutate(Lipotoxicity = replace(Lipotoxicity, Lipotoxicity==0, NA))%>%
  mutate(Glucose = replace(Glucose, Glucose==0, NA)) %>%
  mutate(BloodPressure = replace(BloodPressure, BloodPressure==0, NA)) %>%
  mutate(SkinThickness = replace(SkinThickness, SkinThickness==0, NA)) %>%
  mutate(Insulin = replace(Insulin, Insulin==0, NA)) %>%
  mutate(BMI = replace(BMI, BMI==0, NA))
  
```

```{r}
db <- 
  db %>%
  rename(
    lip = Lipotoxicity,
    glu = Glucose,
    bp = BloodPressure,
    st = SkinThickness,
    ins = Insulin,
    bmi = BMI,
    dpf = DiabetesPedigreeFunction,
    age = Age,
    outcome = Outcome
    )
```


```{r}
db$age_group <- ifelse(db$age < 30, "20-29", 
                         ifelse(db$age >= 30 & db$age < 40, "30-39",
                                ifelse(db$age >= 40 & db$age < 50, "40-49", "50+")))

```

<br>

## Histogram of Diabetes outcome by age groups

```{r}
db <-
  db %>%
  mutate(outcome_f = factor(outcome))

p0 <- 
  db %>%
  ggplot()+
  geom_bar(mapping = aes(x = age_group, fill = outcome_f, stat="identity"))+
  labs(title = "Diabetes outcome by age groups")

ggplotly(p0)

```

<br>

## Histogram of Diabetes Outcome by Diabetes Pedigree Function 

```{r}
p1 <- ggplot(db, aes(x=dpf, color=outcome_f, fill=outcome_f)) +
  scale_fill_manual(values = c("skyblue", "pink")) +
  scale_colour_manual(values = c("black", "black")) +
  geom_histogram()
ggplotly(p1)

```

<br>

## Mean Values of Each Risk Factors by Outcome Group
- Glucose Level and Insulin shows significant difference in mean values by outcome group.

```{r}
db_ot <- 
  db %>% 
  group_by(outcome)
  
db_table <-
  db_ot %>%
  summarise(
    DiabetesPedigreeFunction=mean(dpf, na.rm = TRUE),
    Lipotoxicity=mean(lip, na.rm = TRUE),
    Glucose=mean(glu, na.rm = TRUE),
    BloodPressur=mean(bp, na.rm = TRUE),
    SkinThickness=mean(st, na.rm = TRUE),
    Insulin=mean(ins, na.rm = TRUE),
    BMI=mean(bmi, na.rm = TRUE),
    Age=mean(age, na.rm = TRUE),
    )

knitr::kable(db_table, align= "c", digits = 4, caption = "Mean of each risk factors by outcome group")
```

<br>

## Scatter Plot Graph {.tabset}

### Figure 1

```{r}
db_ot %>% 
  plot_ly(x = ~dpf, y = ~glu, 
          type = 'scatter', mode = 'markers', color = ~outcome_f,
          marker = list(opacity=1.0),
          hoverinfo = 'text',
          text = ~paste( paste("Diabetes Outcome:", outcome_f, sep=""), paste(" Diabetes Pedigree Function: ", dpf, sep="") , paste(" Glucose Level: ", glu, sep=""), sep = "<br>")) %>%
  layout(title = "Glucose Level vs Diabetes Pedigree Function by oucome",
                  yaxis = list(title = "Glucose Level"), xaxis = list(title = "Diabetes Pedigree Function"),
         hovermode = "compare")
```

<br>

### Figure 2

```{r}
db_ot %>% 
  plot_ly(x = ~dpf, y = ~ins, 
          type = 'scatter', mode = 'markers', color = ~outcome_f,
          marker = list(opacity=1.0),
          hoverinfo = 'text',
          text = ~paste( paste("Diabetes Outcome:", outcome_f, sep=""), paste(" Diabetes Pedigree Function: ", dpf, sep="") , paste(" Insulin Level: ", ins, sep=""), sep = "<br>")) %>%
  layout(title = "Insulin Level vs Diabetes Pedigree Function by oucome",
                  yaxis = list(title = "Insulin Level"), xaxis = list(title = "Diabetes Pedigree Function"),
         hovermode = "compare")
```

{-}

<br>

## Scatter Plot Trend {.tabset}

### Glucose Level vs Diabetes Pedigree Function by outcome

```{r}
p_glu <- ggplot(db, aes(x=dpf, y=glu)) + geom_point() + geom_smooth () 
ggplotly(p_glu)
```

<br>

### Insulin Level vs Diabetes Pedigree Function by outcome

```{r}
p_ins <- ggplot(db, aes(x=dpf, y=ins)) + geom_point() + geom_smooth () 
ggplotly(p_ins)
```

{-}